Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-25 paper.tex
use E for environment

Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
machine and Fig.~\ref{fig:state} gives a grammar for machine states
$s$, including a definition of values $v$ and value typing. 
%Fig.~\ref{fig:state} gives a grammar for machine states
%$s$, including a definition of values and value typing and
%Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
%machine.
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machine involves two kinds of transitions, (1) those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation, and (2) those that return a value to the
current continuation and possibly perform a computation.
Corresponding to (1) and (2), the machine state is either in an
$\sOOinspect{e}{\rho}{\kappa}$ or $\sOOreturn{v}{\kappa}$ configuration, 
respectively. Additionally, there is the $\sOOhalt{o}$ configuration, where the 
machine halts with an observation ($o$). Observations include all value 
constructors and blame. The function converting values to 
observations ($observe(v) = o$) is defined in the natural way.
%Let $s$ range over machine states. A state is either looking at an 
%expression to decide what to do next, returning a value to a continuation, or 
%halting with an observation.

@@

\begin{figure}
  Machine state and other runtime data structures
  \[
  \begin{array}{lrcl}
  \stxrule{v}{Values}{
    \vOOtt{} \mid
    \vOOfun{\rho}{x}{e} \mid
    \vOOcons{v}{v} \mid
    \vOOinl{v} \mid
    \vOOinr{v} \mid   
    \vOOcast{v}{c}
  }
  \stxrule{c}{Casts}{
    \cOOcast{T}{l}{T}
  }
  \stxrule{I}{Injectable types (\lazyD)}{
    P
  }
  \stxrule{I}{Injectable types (\lazyUD)}{
    \POOunit \mid
    \POOfun{\star}{\star} \mid
    \star \times \star \mid
    \star + \star
  }
  \stxrule{o}{Observations}{
    \oOOinj \mid
    \oOOsole \mid
    \oOOfun \mid
    \oOOcons \mid
    \oOOinl \mid
    \oOOinr \mid
    \oOOblame{l}
  }
  \stxrule{r}{Cast results}{
    \rOOsucc{v} \mid
    \rOOfail{l}
  }
  \stxrule{s}{States}{
    \sOOinspect{e}{\rho}{\kappa} \mid{}
    \sOOreturn{v}{\kappa} \mid{}
    \sOOhalt{o}
  }
  \stxrule{\kappa}{Continuations}{
    \kOOmt \mid
    \kOOconsI{T}{T}{e}{\rho}{\kappa} \mid
    \kOOconsII{T}{T}{v}{\kappa} \mid
    \kOOinl{T}{T}{\kappa}
  }
  \stxrulecont{
    \kOOinr{T}{T}{\kappa} \mid
    \kOOappI{e}{\rho}{\kappa} \mid
    \kOOappII{v}{\kappa} \mid
    \kOOcar{\kappa}
  }
  \stxrulecont{ 
    \kOOcdr{\kappa}\mid
    \kOOcaseI{e}{e}{\rho}{\kappa}\mid
    \kOOcaseII{v}{e}{\rho}{\kappa}
  }
  \stxrulecont{
    \kOOcaseIII{v}{v}{\kappa} \mid
    \kOOcast{c}{\kappa}
  }
  \end{array}
  \]

@@

  Reduction \fbox{$\judgeCreduce{s}{s}$}
  \[
  \begin{array}{rclr}
    \hiredrule{
    \sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\rho}{\kappa}
  }{
    \sOOinspect{e}{\rho}{
      \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa}
%     \langle\cOOcast{T_1}{l}{T_2}\rangle\kappa
    }
  }{}
  \redrule{
    \sOOinspect{\eOOvar{x}}{\rho}{\kappa}
  }{  
    \sOOreturn{\rho(x)}{\kappa}
  }{}
  \redrule{
    \sOOinspect{\eOOsole}{\rho}{\kappa}
  }{
    \sOOreturn{\vOOtt}{\kappa}
  }{}
  \redrule{
    \sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\rho}{\kappa}
  }{
    \sOOreturn{\vOOfun{\rho}{x}{e}}{\kappa}
  }{}
  \redrule{
    \sOOinspect{\eOOcons{T_1}{T_2}{e_1}{e_2}}{\rho}{\kappa}
  }{
    \sOOinspect{e_1}{\rho}{\kOOconsI{T_1}{T_2}{e_2}{\rho}{\kappa}}
  }{}
  \redrule{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\rho}{\kappa}
  }{
    \sOOinspect{e}{\rho}{\kOOinl{T_1}{T_2}{\kappa}}
  }{}
  \redrule{
  \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\rho}{\kappa}
  }{
  \sOOinspect{e}{\rho}{\kOOinr{T_1}{T_2}{\kappa}}
  }{}
  \redrule{
    \sOOinspect{\eOOapp{e_1}{e_2}}{\rho}{\kappa}
  }{
\sOOinspect{e_1}{\rho}{\kOOappI{e_2}{\rho}{\kappa}}}{}

\redrule{
\sOOinspect{\eOOcar{e}}{\rho}{\kappa}}{
\sOOinspect{e}{\rho}{\kOOcar{\kappa}}}{}

\redrule{
  \sOOinspect{\eOOcdr{e}}{\rho}{\kappa}}{
  \sOOinspect{e}{\rho}{\kOOcdr{\kappa}}}{}

\redrule{
\sOOinspect{\eOOcase{e_1}{e_2}{e_3}}{\rho}{\kappa}}{
\sOOinspect{e_1}{\rho}{\kOOcaseI{e_2}{e_3}{\rho}{\kappa}}}{}

\redrule{
\sOOinspect{\eOOblame{l}}{\rho}{\kappa}}{
\sOOhalt{(\oOOblame{l})}}{}

\redrule{
\sOOreturn{v_1}{\kOOconsI{T_1}{T_2}{e_2}{\rho}{\kappa}}}{
\sOOinspect{e_2}{\rho}{\kOOconsII{T_1}{T_2}{v_1}{\kappa}}}{}

@@

\redrule{
\sOOreturn{v_1}{\kOOappI{e_2}{\rho}{\kappa}}}{
\sOOinspect{e_2}{\rho}{\kOOappII{v_1}{\kappa}}}{}

\redrule{
\sOOreturn{v_2}{\kOOappII{(\vOOfun{\rho}{x}{e})}{\kappa}}}{
\sOOinspect{e}{\rho[x:=v_2]}{\kappa}}{}
  \hiredrule{
    \sOOreturn{v_1}{
      \kOOappII{\vOOcast{v_2}{
          \cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}
      }}{\kappa}
%     (\mathtt{app_2} \; \vOOcast{v_2}{
%       \cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}
%     } \; \kappa)
    }
  }{
    \sOOreturn{v_1}{
      \kOOcast{\cOOcast{T_3}{l}{T_1}}{
        \kOOappII{v_2}{
          \kOOcast{\cOOcast{T_2}{l}{T_4}}{\kappa}
        }
      }
%     \langle\rangle
%     (\mathtt{app_2} \; v_2 \; 
%     \langle\cOOcast{T_2}{l}{T_4}\rangle \kappa}
    }
  }{}
  \redrule{
  \sOOreturn{
    \vOOcons{v_1}{v_2}
  }{\kOOcar{\kappa}}
  }{
  \sOOreturn{v_1}{\kappa}
  }{}
  \hiredrule{
    \sOOreturn{
      \vOOcast{v}{\cOOcast{\POOprod{T_1}{T_2}}{l}{
          \POOprod{T_3}{T_4}}}
    }{\mathtt{fst} \; \kappa}
  }{
    \sOOreturn{v}{
      \kOOcar{
        \kOOcast{\cOOcast{T_1}{l}{T_3}}{\kappa}
      }}
  }{}
  \redrule{
  \sOOreturn{
  \vOOcons{v_1}{v_2}
  }{\kOOcdr{\kappa}}
  }{
  \sOOreturn{v_2}{\kappa}
  }{}

@@

\redrule{
\sOOreturn{v_1}{\kOOcaseI{e_2}{e_3}{\rho}{\kappa}}}{
\sOOinspect{e_2}{\rho}{\kOOcaseII{v_1}{e_3}{\rho}{\kappa}}}{}

\redrule{
\sOOreturn{v_2}{\kOOcaseII{v_1}{e_3}{\rho}{\kappa}}}{
\sOOinspect{e_3}{\rho}{
  \kOOcaseIII{v_1}{v_2}{\kappa}
}}{}

@@

%Let $\kappa$ range over continuations. $\mathtt{stop}$ is the top 
%continuation. The remaining continuations correspond to expressions. For 
%example, $\kOOconsI{T_1}{T_2}{e}{\rho}{\kappa}$ is the continuation waiting 
%for 
%the first part of the pair. And the last continuation, $\kOOcast{c}{\kappa}$ 
%is 
%to cast the value before returning to $\kappa$.

@@

\begin{definition}[Values and cast results for the \effCEK{C} machine] 
\label{def:values-effCEK}
  \[
  \begin{array}{lrcl}
  \stxrule{v}{Values}{
    \hcvOOtt \mid
    \hcvOOfun{c}{\rho}{x}{e}{c} \mid
    \hcvOOcons{v}{c}{v}{c} \mid
    \hcvOOinl{v}{c} \mid
    \hcvOOinr{v}{c} \mid
    \hcvOOinj{I}{v}
  }
\stxrule{r}{Cast results}{
  \rOOsucc{v} \mid
  \rOOfail{l}
}
  \end{array}
  \]
\end{definition}

@@

Next we define the \emph{Lazy D Cast ADT}, which capture the further
requirements that are needed for the theorem that establishes
equivalence to \ineffCEKD{}. Property (1) below states that $id(T)$
acts like the identity function. Property (2) states that $seq(c,c)$
acts like a sequence of casts. Properties (3) through (11) state that
$cast(T_1,l,T_2)$ act like the Lazy D $applyCast$
(Fig.~\ref{fig:applyCast-D-C}).

\begin{definition}[Lazy D Cast ADT]
  \label{def:surely-lazyd}
  A  Cast is a Lazy D Cast if:
  \begin{enumerate}
  \item if $v : T$, then $applyCast(v,id(T)) = \mathtt{succ} \; v $
  \item if $\judgeType{v}{T_1}$,
    $ \judgeTypeFT{c_1}{T_1}{T_2}$, and
    $ \judgeTypeFT{c_2}{T_2}{T_3}$,\\
    then $applyCast(v,seq(c_1,c_2)) = 
    applyCast(v,c_1) >>= \lambda v.applyCast(v,c_2)$\\
    where 
    \[
    \begin{array}{rcl}
      \rOOsucc{v} >>= f & = & f(v) \\
      \rOOfail{l} >>= f & = & \rOOfail{l}
    \end{array}
    \]
  \item if $v : T_1$ and $T_1 \not\smile T_2$,
    then $applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
  \item if $v : \star$, 
    then $applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
  \item if $v : P$,
    then $applyCast(\hcvOOinj{P}{v},cast(\star,l,P')) 
    = applyCast(v,cast(P,l,P')) $
  \item if $v : P$,
    then $applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
  \item if $v : \POOunit$,
    then $applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
  \item if $(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}) : \POOfun{T_1}{T_2}$,
    then\\
    $ 
    applyCast(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}, 
    cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
    = 
    \rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\rho}{x}{b}{seq(c_2,cast(T_2,l,T_4))})}$
  \item if $(\hcvOOcons{v_1}{c_1}{v_2}{c_2}) : \POOprod{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
    \times 
    T_4))$\\
    $ = 
    \rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
    $ 
  \item if $(\hcvOOinl{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
  \item if $(\hcvOOinr{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\$
    applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
  \end{enumerate}
\end{definition}

@@

\begin{figure}
  Syntax
  \[
  \begin{array}{lrcl}
% \stxrule{r}{cast results}{
%   \rOOsucc{v} \mid
%   \rOOfail{l}
% }
% \stxrule{s}{states}{
%   \sOOinspect{e}{\rho}{\kappa} \mid{}
%   \sOOreturn{v}{\kappa} \mid{}
%   \sOOhalt{o}
% }
  \stxrule{\kappa}{Continuation}{
    \kOOcast{c}{k}
  }
  \stxrule{k}{Pre-continuations}{
    \kOOmt \mid{}
    \kOOconsI{T}{T}{e}{\rho}{\kappa} \mid
    \kOOconsII{T}{T}{v}{\kappa} \mid
    \kOOinl{T}{T}{\kappa}
  }
  \stxrulecont{
  	\kOOinr{T}{T}{\kappa} \mid
    \kOOappI{e}{\rho}{\kappa} \mid
    \kOOappII{v}{\kappa} \mid
    \kOOcar{\kappa} \mid
    \kOOcdr{\kappa}
  }
  \stxrulecont{
    \kOOcaseI{e}{e}{\rho}{\kappa} \mid
    \kOOcaseII{v}{e}{\rho}{\kappa} \mid
    \kOOcaseIII{v}{v}{\kappa}
  }
  \end{array}
  \]

@@

  Reduction \fbox{$\judgeSreduce{C}{s}{s}$}
  \[
  \begin{array}{rclr}
  \hiredruleS{
    \sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\rho}{\kappa}
  }{
    \sOOinspect{e}{\rho}{ext(cast(T_1,l,T_2),\kappa)}
  }{}
  \redruleS{
    \sOOinspect{\eOOvar{x}}{\rho}{\kappa}
  }{  
    \sOOreturn{\rho(x)}{\kappa}
  }{}
  \redruleS{
    \sOOinspect{\eOOsole}{\rho}{\kappa}
  }{
    \sOOreturn{\hcvOOtt}{\kappa}
  }{}
  \hiredruleS{
    \sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\rho}{\kappa}
  }{
    \sOOreturn{\hcvOOfun{id(T_1)}{\rho}{x}{e}{id(T_2)}}{\kappa}
  }{}
  \redruleS{
    \sOOinspect{\eOOcons{T_1}{T_2}{e_1}{e_2}}{\rho}{\kappa}
  }{
    \sOOinspect{e_1}{\rho}{
      wrap(\kOOconsI{T_1}{T_2}{e_2}{\rho}{\kappa})
    }
  }{}
  \redruleS{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\rho}{\kappa}
  }{
    \sOOinspect{e}{\rho}{wrap(\kOOinl{T_1}{T_2}{\kappa})}
  }{}
  \redruleS{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\rho}{\kappa}
  }{
    \sOOinspect{e}{\rho}{wrap(\kOOinr{T_1}{T_2}{\kappa})}
  }{}
  \redruleS{
    \sOOinspect{\eOOapp{e_1}{e_2}}{\rho}{\kappa}
  }{
    \sOOinspect{e_1}{\rho}{wrap(\kOOappI{e_2}{\rho}{\kappa})}}{}

  \redruleS{
    \sOOinspect{\eOOcar{e})}{\rho}{\kappa}}{
    \sOOinspect{e}{\rho}{wrap(\kOOcar{\kappa})}}{}

  \redruleS{
    \sOOinspect{\eOOcdr{e}}{\rho}{\kappa}}{
    \sOOinspect{e}{\rho}{wrap(\kOOcdr{\kappa})}}{}

>>>>

Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
machine and Fig.~\ref{fig:state} gives a grammar for machine states
$s$, including a definition of values $v$ and value typing. 
%Fig.~\ref{fig:state} gives a grammar for machine states
%$s$, including a definition of values and value typing and
%Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
%machine.
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machine involves two kinds of transitions, (1) those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation, and (2) those that return a value to the
current continuation and possibly perform a computation.
Corresponding to (1) and (2), the machine state is either in an
$\sOOinspect{e}{\mathcal{E}}{\kappa}$ or $\sOOreturn{v}{\kappa}$ configuration, 
respectively. Additionally, there is the $\sOOhalt{o}$ configuration, where the 
machine halts with an observation ($o$). Observations include all value 
constructors and blame. The function converting values to 
observations ($observe(v) = o$) is defined in the natural way.
%Let $s$ range over machine states. A state is either looking at an 
%expression to decide what to do next, returning a value to a continuation, or 
%halting with an observation.

@@

\begin{figure}
  Machine state and other runtime data structures
  \[
  \begin{array}{lrcl}
  \stxrule{v}{Values}{
    \vOOtt{} \mid
    \vOOfun{\mathcal{E}}{x}{e} \mid
    \vOOcons{v}{v} \mid
    \vOOinl{v} \mid
    \vOOinr{v} \mid   
    \vOOcast{v}{c}
  }
  \stxrule{c}{Casts}{
    \cOOcast{T}{l}{T}
  }
  \stxrule{I}{Injectable types (\lazyD)}{
    P
  }
  \stxrule{I}{Injectable types (\lazyUD)}{
    \POOunit \mid
    \POOfun{\star}{\star} \mid
    \star \times \star \mid
    \star + \star
  }
  \stxrule{o}{Observations}{
    \oOOinj \mid
    \oOOsole \mid
    \oOOfun \mid
    \oOOcons \mid
    \oOOinl \mid
    \oOOinr \mid
    \oOOblame{l}
  }
  \stxrule{r}{Cast results}{
    \rOOsucc{v} \mid
    \rOOfail{l}
  }
  \stxrule{s}{States}{
    \sOOinspect{e}{\mathcal{E}}{\kappa} \mid{}
    \sOOreturn{v}{\kappa} \mid{}
    \sOOhalt{o}
  }
  \stxrule{\kappa}{Continuations}{
    \kOOmt \mid
    \kOOconsI{T}{T}{e}{\mathcal{E}}{\kappa} \mid
    \kOOconsII{T}{T}{v}{\kappa} \mid
    \kOOinl{T}{T}{\kappa}
  }
  \stxrulecont{
    \kOOinr{T}{T}{\kappa} \mid
    \kOOappI{e}{\mathcal{E}}{\kappa} \mid
    \kOOappII{v}{\kappa} \mid
    \kOOcar{\kappa}
  }
  \stxrulecont{ 
    \kOOcdr{\kappa}\mid
    \kOOcaseI{e}{e}{\mathcal{E}}{\kappa}\mid
    \kOOcaseII{v}{e}{\mathcal{E}}{\kappa}
  }
  \stxrulecont{
    \kOOcaseIII{v}{v}{\kappa} \mid
    \kOOcast{c}{\kappa}
  }
  \end{array}
  \]

@@

  Reduction \fbox{$\judgeCreduce{s}{s}$}
  \[
  \begin{array}{rclr}
    \hiredrule{
    \sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{
      \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa}
%     \langle\cOOcast{T_1}{l}{T_2}\rangle\kappa
    }
  }{}
  \redrule{
    \sOOinspect{\eOOvar{x}}{\mathcal{E}}{\kappa}
  }{  
    \sOOreturn{\mathcal{E}(x)}{\kappa}
  }{}
  \redrule{
    \sOOinspect{\eOOsole}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\vOOtt}{\kappa}
  }{}
  \redrule{
    \sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\vOOfun{\mathcal{E}}{x}{e}}{\kappa}
  }{}
  \redrule{
    \sOOinspect{\eOOcons{T_1}{T_2}{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e_1}{\mathcal{E}}{\kOOconsI{T_1}{T_2}{e_2}{\mathcal{E}}{\kappa}}
  }{}
  \redrule{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{\kOOinl{T_1}{T_2}{\kappa}}
  }{}
  \redrule{
  \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
  \sOOinspect{e}{\mathcal{E}}{\kOOinr{T_1}{T_2}{\kappa}}
  }{}
  \redrule{
    \sOOinspect{\eOOapp{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
\sOOinspect{e_1}{\mathcal{E}}{\kOOappI{e_2}{\mathcal{E}}{\kappa}}}{}

\redrule{
\sOOinspect{\eOOcar{e}}{\mathcal{E}}{\kappa}}{
\sOOinspect{e}{\mathcal{E}}{\kOOcar{\kappa}}}{}

\redrule{
  \sOOinspect{\eOOcdr{e}}{\mathcal{E}}{\kappa}}{
  \sOOinspect{e}{\mathcal{E}}{\kOOcdr{\kappa}}}{}

\redrule{
\sOOinspect{\eOOcase{e_1}{e_2}{e_3}}{\mathcal{E}}{\kappa}}{
\sOOinspect{e_1}{\mathcal{E}}{\kOOcaseI{e_2}{e_3}{\mathcal{E}}{\kappa}}}{}

\redrule{
\sOOinspect{\eOOblame{l}}{\mathcal{E}}{\kappa}}{
\sOOhalt{(\oOOblame{l})}}{}

\redrule{
\sOOreturn{v_1}{\kOOconsI{T_1}{T_2}{e_2}{\mathcal{E}}{\kappa}}}{
\sOOinspect{e_2}{\mathcal{E}}{\kOOconsII{T_1}{T_2}{v_1}{\kappa}}}{}

@@

\redrule{
\sOOreturn{v_1}{\kOOappI{e_2}{\mathcal{E}}{\kappa}}}{
\sOOinspect{e_2}{\mathcal{E}}{\kOOappII{v_1}{\kappa}}}{}

\redrule{
\sOOreturn{v_2}{\kOOappII{(\vOOfun{\mathcal{E}}{x}{e})}{\kappa}}}{
\sOOinspect{e}{\mathcal{E}[x:=v_2]}{\kappa}}{}
  \hiredrule{
    \sOOreturn{v_1}{
      \kOOappII{\vOOcast{v_2}{
          \cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}
      }}{\kappa}
%     (\mathtt{app_2} \; \vOOcast{v_2}{
%       \cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}
%     } \; \kappa)
    }
  }{
    \sOOreturn{v_1}{
      \kOOcast{\cOOcast{T_3}{l}{T_1}}{
        \kOOappII{v_2}{
          \kOOcast{\cOOcast{T_2}{l}{T_4}}{\kappa}
        }
      }
%     \langle\rangle
%     (\mathtt{app_2} \; v_2 \; 
%     \langle\cOOcast{T_2}{l}{T_4}\rangle \kappa}
    }
  }{}
  \redrule{
  \sOOreturn{
    \vOOcons{v_1}{v_2}
  }{\kOOcar{\kappa}}
  }{
  \sOOreturn{v_1}{\kappa}
  }{}
  \hiredrule{
    \sOOreturn{
      \vOOcast{v}{\cOOcast{\POOprod{T_1}{T_2}}{l}{
          \POOprod{T_3}{T_4}}}
    }{\mathtt{fst} \; \kappa}
  }{
    \sOOreturn{v}{
      \kOOcar{
        \kOOcast{\cOOcast{T_1}{l}{T_3}}{\kappa}
      }}
  }{}
  \redrule{
  \sOOreturn{
  \vOOcons{v_1}{v_2}
  }{\kOOcdr{\kappa}}
  }{
  \sOOreturn{v_2}{\kappa}
  }{}

@@

\redrule{
\sOOreturn{v_1}{\kOOcaseI{e_2}{e_3}{\mathcal{E}}{\kappa}}}{
\sOOinspect{e_2}{\mathcal{E}}{\kOOcaseII{v_1}{e_3}{\mathcal{E}}{\kappa}}}{}

\redrule{
\sOOreturn{v_2}{\kOOcaseII{v_1}{e_3}{\mathcal{E}}{\kappa}}}{
\sOOinspect{e_3}{\mathcal{E}}{
  \kOOcaseIII{v_1}{v_2}{\kappa}
}}{}

@@

%Let $\kappa$ range over continuations. $\mathtt{stop}$ is the top 
%continuation. The remaining continuations correspond to expressions. For 
%example, $\kOOconsI{T_1}{T_2}{e}{\mathcal{E}}{\kappa}$ is the continuation 
%waiting 
%for 
%the first part of the pair. And the last continuation, $\kOOcast{c}{\kappa}$ 
%is 
%to cast the value before returning to $\kappa$.

@@

\begin{definition}[Values and cast results for the \effCEK{C} machine] 
\label{def:values-effCEK}
  \[
  \begin{array}{lrcl}
  \stxrule{v}{Values}{
    \hcvOOtt \mid
    \hcvOOfun{c}{\mathcal{E}}{x}{e}{c} \mid
    \hcvOOcons{v}{c}{v}{c} \mid
    \hcvOOinl{v}{c} \mid
    \hcvOOinr{v}{c} \mid
    \hcvOOinj{I}{v}
  }
\stxrule{r}{Cast results}{
  \rOOsucc{v} \mid
  \rOOfail{l}
}
  \end{array}
  \]
\end{definition}

@@

Next we define the \emph{Lazy D Cast ADT}, which capture the further
requirements that are needed for the theorem that establishes
equivalence to \ineffCEKD{}. Property (1) below states that $id(T)$
acts like the identity function. Property (2) states that $seq(c,c)$
acts like a sequence of casts. Properties (3) through (11) state that
$cast(T_1,l,T_2)$ act like the Lazy D $applyCast$ for \ineffCEK\
(Fig.~\ref{fig:applyCast-D-C}).

\begin{definition}[Lazy D Cast ADT]
  \label{def:surely-lazyd}
  A  Cast is a Lazy D Cast if:
  \begin{enumerate}
  \item If $v : T$ then $applyCast(v,id(T)) = \mathtt{succ} \; v $
  \item If $\judgeType{v}{T_1}$ and
    $ \judgeTypeFT{c_1}{T_1}{T_2}$ and
    $ \judgeTypeFT{c_2}{T_2}{T_3}$\\
    then $applyCast(v,seq(c_1,c_2)) = 
    applyCast(v,c_1) >>= \lambda v'.applyCast(v',c_2)$\\
    where 
    \[
    \begin{array}{rcl}
      \rOOsucc{v} >>= f & = & f(v) \\
      \rOOfail{l} >>= f & = & \rOOfail{l}
    \end{array}
    \]
  \item If $v : T_1$ and $T_1 \not\smile T_2$,
    then $applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
  \item If $v : \star$,
    then $applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
  \item If $v : P$,
    then $applyCast(\hcvOOinj{P}{v},cast(\star,l,P')) 
    = applyCast(v,cast(P,l,P')) $
  \item If $v : P$,
    then $applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
  \item If $v : \POOunit$,
    then $applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
  \item If $(\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2}) : \POOfun{T_1}{T_2}$,
    then\\
    $ 
    applyCast(\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2}, 
    cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
    = 
    \rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\mathcal{E}}{x}{b}{seq(c_2,cast(T_2,l,T_4))})}$
  \item If $(\hcvOOcons{v_1}{c_1}{v_2}{c_2}) : \POOprod{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
    \times 
    T_4))$\\
    $ = 
    \rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
    $ 
  \item If $(\hcvOOinl{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
  \item If $(\hcvOOinr{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\$
    applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
  \end{enumerate}
\end{definition}

@@

\begin{figure}
  Syntax
  \[
  \begin{array}{lrcl}
% \stxrule{r}{cast results}{
%   \rOOsucc{v} \mid
%   \rOOfail{l}
% }
% \stxrule{s}{states}{
%   \sOOinspect{e}{\mathcal{E}}{\kappa} \mid{}
%   \sOOreturn{v}{\kappa} \mid{}
%   \sOOhalt{o}
% }
  \stxrule{\kappa}{Continuation}{
    \kOOcast{c}{k}
  }
  \stxrule{k}{Pre-continuations}{
    \kOOmt \mid{}
    \kOOconsI{T}{T}{e}{\mathcal{E}}{\kappa} \mid
    \kOOconsII{T}{T}{v}{\kappa} \mid
    \kOOinl{T}{T}{\kappa}
  }
  \stxrulecont{
  	\kOOinr{T}{T}{\kappa} \mid
    \kOOappI{e}{\mathcal{E}}{\kappa} \mid
    \kOOappII{v}{\kappa} \mid
    \kOOcar{\kappa} \mid
    \kOOcdr{\kappa}
  }
  \stxrulecont{
    \kOOcaseI{e}{e}{\mathcal{E}}{\kappa} \mid
    \kOOcaseII{v}{e}{\mathcal{E}}{\kappa} \mid
    \kOOcaseIII{v}{v}{\kappa}
  }
  \end{array}
  \]

@@

  Reduction \fbox{$\judgeSreduce{C}{s}{s}$}
  \[
  \begin{array}{rclr}
  \hiredruleS{
    \sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{ext(cast(T_1,l,T_2),\kappa)}
  }{}
  \redruleS{
    \sOOinspect{\eOOvar{x}}{\mathcal{E}}{\kappa}
  }{  
    \sOOreturn{\mathcal{E}(x)}{\kappa}
  }{}
  \redruleS{
    \sOOinspect{\eOOsole}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\hcvOOtt}{\kappa}
  }{}
  \hiredruleS{
    \sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\hcvOOfun{id(T_1)}{\mathcal{E}}{x}{e}{id(T_2)}}{\kappa}
  }{}
  \redruleS{
    \sOOinspect{\eOOcons{T_1}{T_2}{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e_1}{\mathcal{E}}{
      wrap(\kOOconsI{T_1}{T_2}{e_2}{\mathcal{E}}{\kappa})
    }
  }{}
  \redruleS{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{wrap(\kOOinl{T_1}{T_2}{\kappa})}
  }{}
  \redruleS{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{wrap(\kOOinr{T_1}{T_2}{\kappa})}
  }{}
  \redruleS{
    \sOOinspect{\eOOapp{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e_1}{\mathcal{E}}{wrap(\kOOappI{e_2}{\mathcal{E}}{\kappa})}}{}

  \redruleS{
    \sOOinspect{\eOOcar{e})}{\mathcal{E}}{\kappa}}{
    \sOOinspect{e}{\mathcal{E}}{wrap(\kOOcar{\kappa})}}{}

  \redruleS{
    \sOOinspect{\eOOcdr{e}}{\mathcal{E}}{\kappa}}{
    \sOOinspect{e}{\mathcal{E}}{wrap(\kOOcdr{\kappa})}}{}
